### 2023-10-15 US/before_body.tex
Initial commit

\noindent \textbf{Keywords:} Uncertainty, Monetary transmission, Non-linearity, Sign restriction, Vector-Autoregression \linebreak
\noindent \textbf{JEL codes:} E32, E52, C32 
\newpage
\bgroup
\large{\textbf{Acknowledgements}}\\
\egroup

I would like to express my gratitude to a number of people who helped me see this project through. First and foremost to my PhD advisor, István Kónya for his supervision and guidance over the years of my PhD; to Tohid Atashbar, my former IMF internship supervisor for his co-authorship on the IMF Working Paper, as well as for his suggestions on this paper; to Gergely Gánics for his suggestions on fine-tuning the methodology used in this paper; to Anh Nguyen and Dániel Baksa for their suggestions on alternative specifications that would support my results; as well as to former colleagues from my IMF internship and current colleagues from academia, who provided insightful comments during seminar presentations.

\newpage 
\bgroup
\hypersetup{linkcolor = black}
\renewcommand{\baselinestretch}{1}\normalsize
\tableofcontents
\newpage
\listoffigures
\egroup
\newpage